---
title: "Association Rules"
author: "Cynthia Mwadime"
date: '2022-10-06'
output: pdf_document
---

```{r setup, include=FALSE}
#specifying global settings to be applied to the whole R Markdown script.
# setting all code chunks as “echo=TRUE” so they're all included in the final rendered version.
knitr::opts_chunk$set(echo = TRUE)
```

# Association Analysis

This section ivolves creation of association rules that allow us to identify relationships between variables in the dataset.

We are tasked with creating association rules that will allow us to identify relationships between variables in the dataset. We have been provided with a dataset that comprises of groups of items that will be associated with others. 
### Importing the arules library

```{r}
# Loading the arules library
library(arules, warn.conflicts = FALSE)
```


# Loading the data 

```{r}
# loading
sm <- read.transactions("http://bit.ly/SupermarketDatasetII", sep = ",")
sm
```

```{r}
# Previewing our first 5 transactionss
class(sm)
inspect(sm[1:5])
```



###Generating a summary of the supermarket dataset
This gives us information on stuff like distribution of the item sets, most purchased items and number of items purchased in each transaction among other things 
```{r }
summary(sm)
```

From the report we can see that the most sold items were mineral water], eggs,sphagetti,french fries and chocolate repectively.



```{r}
# Let's see transacations ranging from 6 to 10 
# percentage of the total transactions 
itemFrequency(sm[, 6:10],type = "absolute")
round(itemFrequency(sm[, 6:10],type = "relative")*100,2)
```
### Displaying items whose relative importance is at least 10% 
```{r}
par(mfrow = c(1, 2))
# plot the frequency of items
itemFrequencyPlot(sm, support = 0.1,col="violet")
```
### Displaying top 10 most common items in the transactions dataset and the 
```{r}
itemFrequencyPlot(sm, topN = 10,col="lightblue")
```
### Building a model based on association rules 
* We'll be using the apriori function 
```{r}
# We use Min Support as 0.001 and confidence as 0.8
rules <- apriori (sm, parameter = list(supp = 0.001, conf = 0.7))
rules
```
We obtained a set of 200 rules.

### Tweking the parameters
Let's now see see happens if we increase the support or lower the confidence level. 

```{r}
# Building a apriori model with Min Support as 0.002 and confidence as 0.6
ruls <- apriori (sm,parameter = list(supp = 0.002, conf = 0.6)) 
ruls
```
This gives us a set of 43 rules which is is not enough

```{r}
# Building the apriori model with Min Support as 0.002 and confidence as 0.6.
ruls2 <- apriori (sm, parameter = list(supp = 0.001, conf = 0.6)) 
ruls2
```
This gives us 545 rules. That we can work with

### Let's now get more information on support, lift and confidence 

```{r}

summary(ruls2)
```

Most rules have 3 and 4 items 

```{r}
# Observing the first 5 rules built in our model
inspect(ruls2[1:5])
```

If someone buys cookies and shallot they are 60% likely to buy low fat yogurt.

```{r}
# Ordering the rules by the level of confidence then looking at the first five rules.
ruls2<-sort(ruls2, by="confidence", decreasing=TRUE)
inspect(ruls2[1:5])
```

The first four rules have a confidence of 100%.

```Were interested in creating an ad relating to the sale of a particular item, we could create a subset of rules concerning this product. This would inform us on what items the customers bought before purchasing our target item. Let's use escalope and see our theory in action```

```{r}
escalope <- subset(rules, subset = rhs %pin% "escalope")
# Then order by confidence
escalope <-sort(escalope, by="count", decreasing=TRUE)

```

```{r}
inspect(escalope)
```
* mushroom cream sauce and pasta were in 19 shopping basckets while frenchfries, mushroom creamsauce and pasta had been in 8 basket therefore it would motivate most mushroom cream sauce and pasta buyers to buy escalope as well were they promoted together.


