---
title: "Fraud Detection"
author: "Cynthia Mwadime"
date: '2022-06-10'
output: pdf_document
---

```{r setup, include=FALSE}
#specifying global settings to be applied to the whole R Markdown script.
# setting all code chunks as “echo=TRUE” so they're all included in the final rendered version.
knitr::opts_chunk$set(echo = TRUE)
```

# Anomaly Detection

## Overview
We are tasked with checking whether there are any anomalies in the given sales dataset for the purpose of  fraud detection.


## Loading the Data and Libraries
```{r}
# Loading tidyverse and anomalize
library(tidyverse)
#devtools::install_github("JesseVent/crypto")
library(dplyr)
library(crypto2)
library(anomalize,warn.conflicts = FALSE)
library(tibbletime)
```

```{r}
# reading the data
sales <- read.csv('http://bit.ly/CarreFourSalesDataset')
View(sales)
```

```{r}
# checking the structure of our data
str(sales)
```
* We have 1000 observations and 2 variables.
* We'll have to change the date datatype

```{r}
# converting variables to our preferred format
sales$Date <- as.Date(sales$Date, "%m/%d/%Y")
```

```{r}
# confirming change
str(sales)
```

### Visualizing our Sales

```{r}
# frequency of sales
hist(sales$Sales,col="lightblue")
```

```{r}
#Checking the distribution over time
library(ggplot2)
ggplot(data = sales, aes(x = Date, y = Sales)) +
      geom_bar(stat = "identity", fill = "lightblue") +
      labs(title = "Sales distribution",
           x = "Date", y = "Sales(ksh)")
```


```{r}
# Ordering the data by Date
sales = sales %>% arrange(Date)
head(sales)
```

```Since our data consists of daily records, let's get the average per day so we have more compactdata to work with``
```{r}
forecast <- aggregate(Sales ~ Date , sales , FUN="mean")
head(forecast)
```

```{r}
# Converting data frame into a tibble time (tbl_time) tbl_time have a time index that contains information about which column should be used for time-based subsetting and other time-based manipulation
pred= tbl_time(forecast, Date)
class(pred)
```

We now use the following functions to detect and visualize anomalies; 


```{r}
pred %>%
    time_decompose(Sales) %>%
    anomalize(remainder) %>%
    time_recompose() %>%
    plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5)
```
### Confirming that there aren't anyanomalies
```{r}
skew <- sum(as.numeric(sales$Class))/nrow(sales)
sprintf('Percentage of fraudulent transactions %f', skew*100)
```

## Conclusion
There were no anomalies detected in the data.
